---
title: "SuperCell quick tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SuperCell quick tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/",
  fig.width = 6,
  fig.height = 5,
  eval = TRUE
)
```

```{r setup, warning=FALSE, message=FALSE}
library(SuperCell)
library(Seurat)
```

## 1. Load example data

```{r load-data}
data("bmcite_HumanHTO7")
bmcite_HumanHTO7
```

## 2. Preprocess modalities

`SCimplify_for_Seurat()` uses existing reductions. Here we compute RNA PCA and ADT PCA.

```{r preprocess, warning=FALSE, message=FALSE}
DefaultAssay(bmcite_HumanHTO7) <- "RNA"
bmcite_HumanHTO7 <- NormalizeData(bmcite_HumanHTO7)
bmcite_HumanHTO7 <- FindVariableFeatures(bmcite_HumanHTO7)
bmcite_HumanHTO7 <- ScaleData(bmcite_HumanHTO7)
bmcite_HumanHTO7 <- RunPCA(bmcite_HumanHTO7, reduction.name = "pca")

DefaultAssay(bmcite_HumanHTO7) <- "ADT"
VariableFeatures(bmcite_HumanHTO7) <- rownames(bmcite_HumanHTO7[["ADT"]])
bmcite_HumanHTO7 <- NormalizeData(bmcite_HumanHTO7, normalization.method = "CLR", margin = 2)
bmcite_HumanHTO7 <- ScaleData(bmcite_HumanHTO7)
bmcite_HumanHTO7 <- RunPCA(bmcite_HumanHTO7, reduction.name = "apca")
```

## 3. Build multimodal metacells

```{r build-metacells, warning=FALSE, message=FALSE}
bm_mc <- SCimplify_for_Seurat(
  seurat = bmcite_HumanHTO7,
  assay = c("RNA", "ADT"),
  reduction = list("pca", "apca"),
  dims = list(1:30, 1:18),
  gamma = 75
)

bm_mc
head(FetchData(bm_mc, c("size", "celltype.l1", "celltype.l1_purity")))
```

The output is a Seurat metacell object. It stores metacell size in `size`, aggregated categorical metadata, purity columns, and run metadata in `bm_mc@misc`.

## 4. Optional semi-supervision

Pass a metadata column to `label`. Missing labels may be `NA`.

```{r semisupervised, warning=FALSE, message=FALSE}
set.seed(1)
bmcite_HumanHTO7$partial_annotation <- bmcite_HumanHTO7$celltype.l2
hide <- sample(colnames(bmcite_HumanHTO7), size = 0.5 * ncol(bmcite_HumanHTO7))
bmcite_HumanHTO7$partial_annotation[hide] <- NA

bm_mc_supervised <- SCimplify_for_Seurat(
  seurat = bmcite_HumanHTO7,
  assay = c("RNA", "ADT"),
  reduction = list("pca", "apca"),
  dims = list(1:30, 1:18),
  label = "partial_annotation",
  gamma = 75
)

bm_mc_supervised
```

## 5. Aggregate expression from existing metacell IDs

Use `MetacellExpression()` directly when cells already have metacell IDs. Output column names preserve the grouping IDs.

```{r metacell-expression}
expr <- MetacellExpression(
  object = bmcite_HumanHTO7,
  assays = "RNA",
  group.by = "partial_annotation",
  slot = "counts",
  return.seurat = FALSE
)[["RNA"]]

head(colnames(expr))
```

## 6. Plot metacells

```{r plot-metacells, warning=FALSE, message=FALSE}
bmcite_HumanHTO7 <- RunUMAP(bmcite_HumanHTO7, reduction = "pca", dims = 1:30)
DimPlotSC(
  seurat = bmcite_HumanHTO7,
  seurat.mc = bm_mc,
  reduction = "umap",
  sc.col = "celltype.l2",
  metacell.col = "celltype.l2"
)
```

## 7. Fragment aggregation for ATAC data

For multiome data, pass final metacell barcodes as `membership` values. Unmatched single-cell barcodes are filtered out.

```{r fragment-example, eval=FALSE}
fragment_membership <- c(
  "AAAC-1" = "Metacell_1",
  "AAAG-1" = "Metacell_2"
)

AggregateFragmentFile(
  input_file = "fragments.tsv.gz",
  membership = fragment_membership,
  output_path = "metacell_fragments",
  nb_cl = 4
)
```
